---
title: "Assignment 2 Part B"
output: pdf_document
date: "2024-11-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ISLR)
library(dplyr)
library(glmnet)
library(caret)
```

```{r}
Carseats_Filtered <- Carseats %>% select("Sales", "Price",
"Advertising","Population","Age","Income","Education")
```

#QB1. Build a decision tree regression model to predict Sales based on all other attributes ("Price", "Advertising", "Population", "Age", "Income" and "Education"). Which attribute is used at the top of the tree (the root node) for splitting? Hint: you can either plot () and text() functions or use the summary() function to see the decision tree rules.

```{r}
# First, let's make sure we have the rpart package
if (!require(rpart)) install.packages("rpart")
library(rpart)

# Build the decision tree model
tree_model <- rpart(Sales ~ ., data = Carseats_Filtered)

# Print the summary of the model
summary(tree_model)

# Plot the tree
plot(tree_model)
text(tree_model, pretty = 0)
```
#QB2. Consider the following input:
#• Sales=9
#• Price=6.54
#• Population=124
#• Advertising=0
#• Age=76
#• Income= 110
#• Education=10
#What will be the estimated Sales for this record using the decision tree model? (15% of total points)

```{r}
# Create a new data frame with the given input
new_data <- data.frame(
  Price = 6.54,
  Population = 124,
  Advertising = 0,
  Age = 76,
  Income = 110,
  Education = 10
)

# Use the tree model to predict Sales for the new data
predicted_sales <- predict(tree_model, newdata = new_data)

# Print the predicted Sales
print(predicted_sales)
```

#QB3. Use the caret function to train a random forest (method=’rf’) for the same dataset. Use the caret default settings. By default, caret will examine the “mtry” values of 2,4, and 6. Recall that mtry is the number of attributes available for splitting at each splitting node. Which mtry value gives the best performance? (Make sure to set the random number generator seed to 123)

```{r}
# Set the random seed for reproducibility
set.seed(123)

# Train the random forest model using caret
rf_model <- train(Sales ~ ., 
                  data = Carseats_Filtered, 
                  method = "rf",
                  trControl = trainControl(method = "cv", number = 5))

# Print the results
print(rf_model)
```

#QB4. Customize the search grid by checking the model’s performance for mtry values of 2, 3 and 5 using 3 repeats of 5-fold cross validation.

```{r}
# Set the random seed for reproducibility
set.seed(123)

# Define the custom search grid
custom_grid <- expand.grid(mtry = c(2, 3, 5))

# Define the control parameters for training
ctrl <- trainControl(method = "repeatedcv",
                     number = 5,
                     repeats = 3)

# Train the random forest model with the custom grid
rf_model_custom <- train(Sales ~ ., 
                         data = Carseats_Filtered, 
                         method = "rf",
                         metric = "RMSE",
                         tuneGrid = custom_grid,
                         trControl = ctrl)

# Print the results
print(rf_model_custom)
```

